Comparison of linear beam theories

### 2024-03-21 article.tex
article v2
Article draft final

\author{Rudi du Toit\corref{cor1}}
\ead{u15000258@tuks.co.za}
\cortext[cor1]{Corresponding author}
\author{Prof N.F.J. van Rensburg}
\address{Department of Mathematics, University of Pretoria, Pretoria, South Africa}

@@

\begin{keyword}
	Beams \sep Cantilever Beams \sep Two-Dimensional Models \sep Three-Dimensional Models \sep Vibration Analysis \sep Modal Analysis \sep Finite Element Method
\end{keyword}

@@

\section{Linear mathematical models}
	\subsection{Three-dimensional theory}
		Consider a vector valued displacement function $u \in \Omega \subset R^3$.

@@

		\subsubsection*{Constitutive Equation}
			\textcolor{red}{************ VOORSTEL om the verkort ************}

			For an isotropic material, Hooke's law can be written in terms of Young's modulus $E$ and Poisson's ratio $\nu$. And if the principal stresses $\sigma_i$ are all non-zero, then Hooke's law can be written in the following form:
			\begin{eqnarray}
				T = \left( \frac{E}{1+\nu} \right)\mathcal{E} + \frac{\nu E}{(1+\nu)(1-2\nu)}\textrm{Tr}(\mathcal{E})I \label{eq:3D_Model:CE}.
			\end{eqnarray} Here $\mathcal{E}$ is the strain tensor and $T$ is the stress tensor.

@@

			\textcolor{red}{************ VOORSTEL ************}

			For an isotropic material, Hooke's law can be written in terms of Young's modulus $E$ and Poisson's ratio $\nu$ as follows.

			\textbf{Hooke's law}
			\begin{eqnarray}
				\mathcal{E} = \left( \frac{1+\nu}{E} \right)T - \frac{\nu}{E}\textrm{Tr}(T)I,  \label{eq:3D_Model:HL}
			\end{eqnarray} The components of the strain tensor $\mathcal{E}$ are defined as
			\begin{eqnarray}
				\varepsilon_{ij} = \frac{1}{2}\left( \frac{\partial u_i}{\partial x_j} + \frac{\partial u_j}{\partial x_i} \right). \label{eq:3D_Model:Strain}
			\end{eqnarray}\label{sym:varepsilonij}

			If the principal stresses $\sigma_i$ are all non-zero, then Hooke's law can be written in an alternative form.

			\textbf{Hooke's law in the alternative form}
			\begin{eqnarray}
				T = \left( \frac{E}{1+\nu} \right)\mathcal{E} + \frac{\nu E}{(1+\nu)(1-2\nu)}\textrm{Tr}(\mathcal{E})I \label{eq:3D_Model:CE}.
			\end{eqnarray}
			With $\textrm{Tr}(\mathcal{E}) = \epsilon_{11} + \epsilon_{12} + \epsilon_{13}$ is the trace of the strain operator $\mathcal{E}$.

		\subsubsection*{Dimensionless form} \label{ssec:3D_Model:DimensionlessForm}
			Let $\ell$ represent a significant dimension (e.g., length) of an elastic body and $G$ the shear modulus. Define dimensionless variables and normalized functions as
			\[\tau = \frac{t}{t_0},\ \xi_i = \frac{x_i}{\ell},\ u^*(\xi,\tau) = \frac{u(x,t)}{\ell},\ Q^{*} = \ell G \kappa^2 Q,\ \sigma_{ij}^*(\xi,\tau) = \frac{\sigma_{ij}(x,t)}{G\kappa^2},\]
			where $\kappa^2$ is a dimensionless constant and $\displaystyle t_0 = \ell\sqrt{\frac{\rho}{G\kappa^2}}$. Let $\displaystyle \gamma= \frac{G\kappa^2}{E}$.

@@

		\subsubsection*{Equations of motion in dimensionless form}\label{sssec:3D_Model:EquationOfMotionDimensionless}
			\begin{eqnarray}
				\partial_t^2 u & = & \textrm{div}T + Q \label{eq:3D_Model:EM-D}
			\end{eqnarray}
			with
			\begin{eqnarray}
				\textrm{div}  T & = &
				\begin{bmatrix}
					\partial_1 \sigma_{11} + \partial_2 \sigma_{12} + \partial_3 \sigma_{13} \\
					\partial_1 \sigma_{21} + \partial_2 \sigma_{22} + \partial_3 \sigma_{23} \\
					\partial_1 \sigma_{31} + \partial_2 \sigma_{32} + \partial_3 \sigma_{33}
				\end{bmatrix}.\label{eq:3D_Model:divT-D}
			\end{eqnarray}

		\subsubsection*{Constitutive equations in dimensionless form}\label{sssec:3D_Model:ConstitutiveEquationsDimensionless}
			\begin{eqnarray}
				T = \frac{1}{\gamma(1+\nu)} \mathcal{E} + \frac{\nu}{\gamma(1+\nu)(1-2\nu)}\textrm{Tr}(\mathcal{E})I \label{eq:3D_Model:CE-D}
			\end{eqnarray}

		Divide the boundary of $\Omega$ into two distinct parts, referred to as $\Sigma$ and $\Gamma$. The following will be considered a model problem for a three-dimensional elastic body (also for the two-dimensional elastic body).
		\begin{enumerate}
			\item[] The equations of motion \eqref{eq:3D_Model:EM-D} and \eqref{eq:3D_Model:divT-D}, and Hooke's law is satisfied in $\Omega$;
			\item[] The displacement $u = u_\Sigma$ is specified on
				$\Sigma$; The traction $Tn = t_\Gamma$ is specified on
				$\Gamma$.
		\end{enumerate} \label{sym:n}\label{sym:SigmaGamma}

		\subsubsection*{Cantilever model}
			Find a vector valued function $u$, satisfying equations
			\eqref{eq:3D_Model:EM-D} to \eqref{eq:3D_Model:CE-D} and the following boundary
			conditions:

			Boundary Conditions:
			\begin{eqnarray*}
				u & = & 0 \quad \textrm{ on } \Sigma\\
				Tn & = & 0 \quad \textrm{ on } \Gamma
			\end{eqnarray*}

			This will be referred to as \textbf{Problem 3D}.

		\subsection*{Variational form}\label{ssec:3D_Model:VariationalForm}
			Define the bilinear forms
			\begin{eqnarray}
				b(u,\phi) = \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
				\textrm{Tr}(\Phi) ~dV \label{eq:3D_Model:Bilinear}
			\end{eqnarray}\label{sym:b}
			and
			\begin{eqnarray}
				c(u,\phi) = \int_\Omega~ (\partial^2_t u) \cdot \phi~dV \label{eq:3D_Model:Bilinear_c}
			\end{eqnarray}
			with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1+\nu)(1-2\nu)}$.

			Also define the integral
			\begin{eqnarray}
				(f,g) &=& \int_{\Omega} f\cdot g \ dV \label{eq:3D_Model:Bilinear_int}
			\end{eqnarray}

			Define the following set of test functions $T(\Omega)$ for Problem 3D:
			\begin{eqnarray*}
				T(\Omega) & = & \left\{ \phi \in C^1(\bar{\Omega}) \ | \ \phi = 0 \ \textrm{ on } \ \Gamma \right\}
			\end{eqnarray*}\label{sym:TOmega}

			Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and
			\begin{align}
				c(u,\phi) = -b(u,\phi) + (Q,\phi), \label{eq:3D_Model:Problem3D1VEq}
			\end{align}
			for all $\phi \in T(\Omega)$

			This will be referred to as \textbf{Problem 3DV}.

@@

			Find a vector valued function $u$, satisfying equations \eqref{eq:2D_Model:EM} to \eqref{eq:2D_Model:CE} and the boundary conditions:

			Boundary Conditions:
			\begin{eqnarray*}
				u & = & 0 \quad \textrm{ on } \Sigma,\\
				Tn & = & 0 \quad \textrm{ on } \Gamma.
			\end{eqnarray*} with $n$ the outward unit normal.

			This will be referred to as \textbf{Problem 2D}.

		\subsubsection*{Variational form}\label{ssec:2D_Model:VariationalForm}
			Define the bilinear forms
			\begin{eqnarray}
				b(u,\phi) = \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
				\textrm{Tr}(\Phi) ~dA \label{eq:2D_Model:Bilinear}
			\end{eqnarray}
			and
			\begin{eqnarray}
				c(u,\phi) = \int_\Omega~ (\partial^2_t u) \cdot \phi~dA \label{eq:2D_Model:Bilinear_c}
			\end{eqnarray}
			with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1-\nu^2)}$.
			Also define the integral
			\begin{eqnarray}
				(f,g) &=& \int_{\Omega} f\cdot g \ dA \label{eq:2D_Model:Bilinear_int}
			\end{eqnarray}

			\subsubsection*{Test functions for Problem 2D}\label{sssec:2D_Model:TestFunctions2D1}
			Define the following set of test functions $T(\Omega)$ for Problem 2D:
			\begin{eqnarray*}
				T(\Omega) & = & \left\{ \phi \in C^1(\bar{\Omega}) \ | \ \phi = 0 \ \textrm{ on } \ \Gamma \right\}. 
			\end{eqnarray*}\label{sym:baromega}

			Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and 
			\begin{align}
				c(u,\phi) = -b(u,\phi) + (Q,\phi) \label{eq:2D_Model:Problem2D1VEq}
			\end{align}
			for all $\phi \in T(\Omega)$. This will be referred to as \textbf{Problem 2DV}.

@@

		\subsection{Models and variational forms}
			\subsubsection*{Bilinear forms and integral}
				For $f,g \in T[0,1]$, define the bilinear forms
				\begin{eqnarray}
					c(f,g) & = & \int_0^1  \partial_t^2 f_1 g_1 + \frac{1}{\alpha}\int_0^1  \partial_t^2 f_2g_2, \label{eq:1D_Model:Bilinear} \\
					b(f,g) & = &   \int_0^1 (f_1'-f_2)(g_1' - g_2) + \frac{1}{\beta}\int_0^1 f_2' g_2', \label{eq:1D_Model:Bilinear_c}
				\end{eqnarray}
				Define the integral
				\begin{eqnarray}
				(f,g)
					&=& \int_{0}^1 fg \label{eq:1D_Model:Bilinear_int}
				\end{eqnarray} for all $f,g \in L^2(0,1).$\\

				\textbf{Remark}\\
				$L^2(a,b)$ is the space of all square integrable functions on the interval $(a,b)$. The inner product is defined by $\displaystyle (f,g) = \int_a^b fg$, and the induced norm $\displaystyle ||f|| = \int_a^b f^2$.\\

@@

				{Boundary Conditions}\\
				\begin{eqnarray*}
					w(0,\cdot) = 0, \ \ &M(0,\cdot) = 0, \label{eq:1D_Model:ProblemT1BC1}\\
					w(1,\cdot) = 0, \ \ &M(1,\cdot) = 0. \label{eq:1D_Model:ProblemT1BC2}
				\end{eqnarray*}

@@

				{Boundary Conditions}\\
				\begin{eqnarray*}
					w(0,\cdot) = 0, \ \ &\phi(0,\cdot) = 0, \label{eq:1D_Model:ProblemT2BC1}\\
					M(1,\cdot) = 0, \ \ &V(1,\cdot) = 0. \label{eq:1D_Model:ProblemT2BC2}
				\end{eqnarray*}

@@

\section{Modal analysis and comparison of linear models}

@@

\section{Cantilever two-dimensional model}\label{ssec:2D_Model:FEM}
Let $\{e_1,e_2\}$ be a right-handed orthonormal basis for $R^2$. Denote the elastic body by $\Omega \in R^2$, with $(0,0)$ as the point of reference. For a rectangular cross-section, the body $\Omega$ can be described as,
\begin{eqnarray*}
	\Omega = \left \{ x \in E_2 \ | \ 0 \leq x_1 \leq 1, \ -\frac{h}{2} \leq x_2 \leq \frac{h}{2} \right \},
\end{eqnarray*} where $\partial \Omega$ denotes the boundary of $\Omega$.

@@

 Using the reference configuration, the variational problem Problem 2DV can be rewritten into a model specific form.

\subsubsection*{Problem 2DV}\label{sssec:2D_Model:Problem2D1VChp5}
Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and 
\begin{align}
	c(u,\phi) = -b(u,\phi) + (Q,\phi) \label{eq:2D_Model:Problem2D1VEqChp5}
\end{align}
for all $\phi \in T(\Omega)$. With the test function space 
\begin{eqnarray*}
	T(\Omega) & = & \left\{ \phi \in C^1(\bar{\Omega}) \ | \ \phi = 0 \ \textrm{ on } \ \Gamma \right\}.
\end{eqnarray*}

The bilinear forms and integral function is defined by
\begin{eqnarray}
	b(u,\phi) & = & \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
	\textrm{Tr}(\Phi) ~dA, \label{eq:2D_Model:BilinearChp5}\\
	c(\partial_t^2 u,\phi) & = & \int_\Omega~ (\partial^2_t u) \cdot \phi~dA, \label{eq:2D_Model:Bilinear_cChp5}\\
	(f,g) &=& \int_{\Omega} f\cdot g \ dA, \label{eq:2D_Model:Bilinear_intChp5}
\end{eqnarray}
with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1-\nu^2)}$.

Using the reference configuration, the constitutive equations and the bilinear form $b$ can be rewritten into a model specific form.

\subsubsection*{Constitutive equations:}
\begin{eqnarray}
	\sigma_{11} & = & \frac{1}{\gamma(1-\nu^2)}(\partial_1 u_1 + \nu \partial_2 u_2) \label{CE1} \\
	\sigma_{22} & = & \frac{1}{\gamma(1-\nu^2)}(\partial_2 u_2 + \nu \partial_1 u_1) \label{CE2} \\
	\sigma_{12} & = & \frac{1}{2\gamma(1+\nu)}(\partial_1 u_2 + \partial_2 u_1) \label{CE3}
\end{eqnarray}

\subsubsection*{Bilinear form:}
\begin{align}
	b(u,\phi) & = & \frac{1}{\gamma(1-\nu^2)}\int_{\Omega} (\partial_1 u_1 \partial_1 \phi_1 + \partial_2 u_2 \partial _2 \phi_2 + \nu\partial_1 u_1 \partial_2\phi_2 + \nu \partial_2 u_2 \partial_1 \phi_1 ) \ dA \nonumber\\
	& + & \frac{1}{2\gamma(1+\nu)}\int_{\Omega} (\partial_1 u_2 \partial_1 \phi_2 + \partial_1 u_2 \partial_2 \phi_1 + \partial_2 u_1 \partial_1\phi_2 + \partial_2 u_1 \partial_2\phi_1) \ dA.
\end{align}

@@

\section{Cantilever three-dimensional model} \label{ssec:3D_Model:FEM}
	Let $\{e_1,e_2,e_3\}$ be a right-handed orthonormal basis for $R^3$. Denote the elastic body as $\Omega \in R^3$ with $(0,0,0)$ the point of reference. For a rectangular cross-section, the body $\Omega$ can be described as
	\begin{eqnarray*}
		\Omega = \left\{ x \in R^3 \ | \ 0 \leq x_1 \leq 1, \ -\frac{h}{2} \leq x_2 \leq \frac{h}{2} , \ -\frac{b}{2} \leq x_3 \leq \frac{b}{2}\right \}
	\end{eqnarray*}
	Let $\partial \Omega$ denote the boundary of the body.

@@

	This general variational form Problem 3DV can now be rewritten as in a model specific form using the reference configuration.

	\subsubsection*{Problem 3DV}
	Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and 
	\begin{align}
		c(u,\phi) = -b(u,\phi) - (Q,\phi) \label{3DB_20}
	\end{align}
	for all $\phi \in T(\Omega)$.\\

	With the test function space
	\begin{eqnarray*}
		T(\Omega) & = & \left\{ \phi \in C(\Omega) \ | \ \phi = 0 \ \textrm{ on } \ \Gamma \right\}.
	\end{eqnarray*}

	The bilinear forms and integral function are defined by
	\begin{eqnarray}
		b(u,\phi) &=& \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
		\textrm{Tr}(\Phi) ~dV,\\
		c(u,\phi) &=& \int_\Omega~ (\partial^2_t u) \cdot \phi~dV, \\
		(f,g) &=& \int_{\Omega} f\cdot g \ dV, \\
	\end{eqnarray}
	with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1+\nu)(1-2\nu)}$.\\

	Using the definition of the reference configuration, the constitutive equations and the bilinear form $b$ can be rewritten in the following model specific forms: 
	\subsubsection*{Constitutive equations}
	\begin{eqnarray*}
		\sigma_{11} & = &  \frac{1}{\gamma(1+\nu)}\partial_1 u_1 + \frac{\nu}{\gamma (1+\nu)(1-2\nu)}(\partial_1 u_1 +  \partial_2 u_2 + \partial_3 u_3)\label{3DB_3}\\
		\sigma_{22} & = &  \frac{1}{\gamma(1+\nu)}\partial_2 u_2 + \frac{\nu}{\gamma (1+\nu)(1-2\nu)}(\partial_1 u_1 +  \partial_2 u_2 + \partial_3 u_3)\label{3DB_4}\\
		\sigma_{33} & = &  \frac{1}{\gamma(1+\nu)}\partial_3 u_3 + \frac{\nu}{\gamma (1+\nu)(1-2\nu)}(\partial_1 u_1 +  \partial_2 u_2 + \partial_3 u_3)\label{3DB_5}\\
		\sigma_{23} & = &  \frac{1}{2\gamma(1+\nu)}(\partial_3 u_2 + \partial_2 u_3)\label{3DB_6}\\
		\sigma_{31} & = &  \frac{1}{2\gamma(1+\nu)}(\partial_3 u_1 + \partial_1 u_3)\label{3DB_7}\\
		\sigma_{12} & = &  \frac{1}{2\gamma(1+\nu)}(\partial_2 u_1 + \partial_1 u_2)\label{3DB_8}
	\end{eqnarray*}

	\subsubsection*{Bilinear Form}
	\begin{align*}
		b(u,\phi)  &=  \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
		\textrm{Tr}(\Phi) ~dV \nonumber\\
		&=  \int_{\Omega} \sigma_{11} \partial_1\phi_1 + \sigma_{12}\partial_1 \phi_2 + \sigma_{13}\partial_1 \phi_3 + \sigma_{21}\partial_2 \phi_1 + \sigma_{22}\partial_2\phi_2 + \sigma_{23}\partial_2 \phi_3 ~dV
	\end{align*}

@@

	\subsubsection*{Problem 3DG}
	Find a function $u^h$ such that for all $t>0$, $u^h \in S^h$ and
	\begin{eqnarray*}
		(u^h, \phi_i) & = & -b(u^h,\phi_i) + (Q^I \cdot \phi_i)
	\end{eqnarray*} for $i = 1,2,...,k$. $Q^I$ is scalar vector obtained after interpolating the function $Q$ over the rectangular grid $\Omega$. i.e. $Q^I_{i,j, h} = Q(x_i,x_j, x_h)$ for $i = 1,2,...,n_1$, $j = 1,2,...,n_2$ and $h = 1,2,...,n_3$.

@@

\section{Validity of the model for a cantilever Timoshenko beam (CHAPTER 4 of dissertation)} \label{sec:validity-of-a-cantilever-timoshenko-beam}

	In this section, the validity of the Timoshenko beam model is investigated. Specifically, the validity of a cantilever Timoshenko beam model, using a cantilever two-dimensional beam model as a reference. This follows what is done in the article \cite{LVV09}. This section discusses the results of the article, focussing on the numerical results. Some of the results are replicated with more significant digits, others are extended to be used in the rest of this chapter.

	The article \cite{LVV09} is titled `Comparison of linear beam theories'. In this article, the authors compare a cantilever Timoshenko beam to a cantilever two-dimensional beam. The authors start by presenting the models. These are the same as Problem T-2 and Problem 2D-1 that are defined in Chapter~1. The authors also look at the existence and uniqueness of solutions, which is covered in Chapter 2 of this dissertation. The authors then calculate and compare the eigenvalues and eigenfunctions, which will be discussed in this section.

	\subsection{The models}
	The cantilever Timoshenko beam model is defined in section \ref{ssec:1D_Model:ModelProblems}, and is referred to as Problem T-2. The cantilever two-dimensional beam model is defined in section \ref{ssec:2D_Model:ModelProblem}, and is referred to as Problem 2D-1.

	Figure \ref{fig:compare:1D+2D} shows the two beams side-by-side.

@@

						%\node at (3.2,-0.2) {$\ell = 1$};
						\end{tikzpicture}
					\end{center}
					\subcaption{Timoshenko Cantilever Beam}
				\end{minipage}
				\begin{minipage}[b]{0.8\linewidth}
					\begin{center}
						\begin{tikzpicture}
							\draw[line width = 0.4mm] (0,0.2) -- (6,0.2);
							\draw[line width = 0.4mm] (0,-0.2) -- (6,-0.2);
							\draw[line width = 0.4mm] (6,-0.2) -- (6,0.2);
							%\draw[line width = 0.4mm] (0,-0.2) -- (0,0.2);

@@

						\end{tikzpicture}
					\end{center}
					\subcaption{Two-Dimensional Cantilever Beam}
				\end{minipage}
			}
		}
	\end{figure}\label{fig:compare:1D+2D}

	In the derivation of the Timoshenko beam model in section \ref{subsec:equationsofmotion+constitutiveequation}, the parameter $\alpha$ is 
	introduced. The paramater is given here again for convenience:
	\begin{eqnarray*}
		\alpha = \frac{A\ell^2}{I}.
	\end{eqnarray*}

>>>>

\author{R. du Toit\corref{cor1}}
\ead{u15000258@tuks.co.za}
\cortext[cor1]{Corresponding author}
\author{N.F.J. van Rensburg}
\address{Department of Mathematics, University of Pretoria, Pretoria, South Africa}

@@

\begin{keyword}
	Beam \sep Cantilever Beam \sep Two-Dimensional Model \sep Three-Dimensional Model \sep Vibration Analysis \sep Modal Analysis \sep Finite Element Method
\end{keyword}

@@

The layout of the paper is as follows:
\begin{itemize}
\item Section \ref{sec:LinearMathematicalModels} introduces the linear mathematical models for the three-dimensional, two-dimensional, and Timoshenko beam models.
\item Section \ref{sec:ModalAnalysis} covers modal analysis, compares linear models, and touches on the models' well-posedness.
\item Section \ref{ssec:2D_Model:FEM} outlines the variational problem for the two-dimensional model and applies the Finite Element Method to derive an eigenvalue problem.
\item Section \ref{ssec:3D_Model:FEM} outlines the variational problem for the three-dimensional model and applies the Finite Element Method to derive an eigenvalue problem.
\item Section \ref{sec:validity-of-a-cantilever-timoshenko-beam} examines the validity of the Timoshenko model for a cantilever beam, drawing on results from \cite{LVV09}.
\item Section \ref{sec:validity-of-a-2d-beam} expands on \cite{LVV09} to assess the validity of a two-dimensional model for a cantilever beam.
\item Section \ref{sec:conclusion} concludes the paper.
\end{itemize}

\section{Linear mathematical models} \label{sec:LinearMathematicalModels}
	\subsection{Three-dimensional theory}
		Consider a vector valued displacement function $u$ defined on $\Omega \subset R^3$.

@@

		\subsubsection*{Constitutive Equation}
			For an isotropic material, Hooke's law can be written in terms of Young's modulus $E$ and Poisson's ratio $\nu$. And if the principal stresses $\sigma_i$ are all non-zero, then Hooke's law can be written in the following form:
			\begin{eqnarray}
				T = \left( \frac{E}{1+\nu} \right)\mathcal{E} + \frac{\nu E}{(1+\nu)(1-2\nu)}\textrm{Tr}(\mathcal{E})I \label{eq:3D_Model:CE}.
			\end{eqnarray} Here $\mathcal{E}$ is the strain tensor and $T$ is the stress tensor.

@@

			Let $\ell$ represent a significant dimension (e.g., length) of an elastic body and $G$ the shear modulus. Define dimensionless variables and normalized functions as
			\[\tau = \frac{t}{t_0},\ \xi_i = \frac{x_i}{\ell},\ u^*(\xi,\tau) = \frac{u(x,t)}{\ell},\ Q^{*} = \ell G \kappa^2 Q,\ \sigma_{ij}^*(\xi,\tau) = \frac{\sigma_{ij}(x,t)}{G\kappa^2},\]
			where $\kappa^2$ is a dimensionless constant and $\displaystyle t_0 = \ell\sqrt{\frac{\rho}{G\kappa^2}}$. Let $\displaystyle \gamma= \frac{G\kappa^2}{E}$.

@@

		\textbf{Equations of motion in dimensionless form}
			\begin{eqnarray}
				\partial_t^2 u & = & \textrm{div}T + Q \label{eq:3D_Model:EM-D}
			\end{eqnarray}
			with
			\begin{eqnarray}
				\textrm{div}  T & = &
				\begin{bmatrix}
					\partial_1 \sigma_{11} + \partial_2 \sigma_{12} + \partial_3 \sigma_{13} \\
					\partial_1 \sigma_{21} + \partial_2 \sigma_{22} + \partial_3 \sigma_{23} \\
					\partial_1 \sigma_{31} + \partial_2 \sigma_{32} + \partial_3 \sigma_{33}
				\end{bmatrix}.\label{eq:3D_Model:divT-D}
			\end{eqnarray}

		\textbf{Constitutive equations in dimensionless form}
			\begin{eqnarray}
				T = \frac{1}{\gamma(1+\nu)} \mathcal{E} + \frac{\nu}{\gamma(1+\nu)(1-2\nu)}\textrm{Tr}(\mathcal{E})I \label{eq:3D_Model:CE-D}
			\end{eqnarray}

		\subsubsection*{Cantilever model}
			Find a vector valued function $u$, satisfying equations
			\eqref{eq:3D_Model:EM-D} to \eqref{eq:3D_Model:CE-D} and the boundary
			conditions $\displaystyle u = 0$ on part of the boundary and $\displaystyle Tn = 0$ on the rest of the boundary. This will be referred to as \textbf{Problem 3D}.

@@

			Find a vector valued function $u$, satisfying equations \eqref{eq:2D_Model:EM} to \eqref{eq:2D_Model:CE} and the boundary conditions $u = 0$ on part of the boundary and $Tn = 0$ on the rest of the boundary. This will be referred to as \textbf{Problem 2D}.

@@

		\textbf{Models and variational forms}
			\subsubsection*{Bilinear forms and integral}
				For $f,g \in T[0,1]$, define the bilinear forms
				\begin{eqnarray}
					c(f,g) & = & \int_0^1  \partial_t^2 f_1 g_1 + \frac{1}{\alpha}\int_0^1  \partial_t^2 f_2g_2, \label{eq:1D_Model:Bilinear} \\
					b(f,g) & = &   \int_0^1 (f_1'-f_2)(g_1' - g_2) + \frac{1}{\beta}\int_0^1 f_2' g_2', \label{eq:1D_Model:Bilinear_c}
				\end{eqnarray}
				Define the integral
				\begin{eqnarray}
				(f,g)
					&=& \int_{0}^1 fg \label{eq:1D_Model:Bilinear_int}
				\end{eqnarray} for all $f,g \in L^2(0,1).$

				\textbf{Remark}
				$L^2(a,b)$ is the space of all square integrable functions on the interval $(a,b)$. The inner product is defined by $\displaystyle (f,g) = \int_a^b fg$, and the induced norm $\displaystyle ||f|| = \int_a^b f^2$.\\

@@

				{Boundary Conditions}
				\begin{eqnarray*}
					w(0,\cdot) = 0, \ \ &M(0,\cdot) = 0, \label{eq:1D_Model:ProblemT1BC1}\\
					w(1,\cdot) = 0, \ \ &M(1,\cdot) = 0. \label{eq:1D_Model:ProblemT1BC2}
				\end{eqnarray*}

@@

				{Boundary Conditions}
				\begin{eqnarray*}
					w(0,\cdot) = 0, \ \ &\phi(0,\cdot) = 0, \label{eq:1D_Model:ProblemT2BC1}\\
					M(1,\cdot) = 0, \ \ &V(1,\cdot) = 0. \label{eq:1D_Model:ProblemT2BC2}
				\end{eqnarray*}

@@

\section{Modal analysis and comparison of linear models}\label{sec:ModalAnalysis}

@@

\section{Cantilever two-dimensional model in variational form}\label{ssec:2D_Model:FEM}
Let $\{e_1,e_2\}$ be a right-handed orthonormal basis for $R^2$. Denote the elastic body by $\Omega \in R^2$, with $(0,0)$ as the point of reference. For a rectangular cross-section, the body $\Omega$ can be described as,
\begin{eqnarray*}
	\Omega = \left \{ x \in E_2 \ | \ 0 \leq x_1 \leq 1, \ -\frac{h}{2} \leq x_2 \leq \frac{h}{2} \right \},
\end{eqnarray*} 
$\partial \Omega$ denotes the boundary of the body. $\Sigma$ is the part of the boundary where $x_1 = 0$ and here the boundary condition is $u = 0$. $\Gamma$ is the remainder of the boundary where $Tn = 0$.

@@

\subsubsection*{Problem 2DV}\label{sssec:2D_Model:Problem2D1VChp5}
The bilinear forms and integral function is defined by
\begin{eqnarray}
	b(u,\phi) & = & \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
	\textrm{Tr}(\Phi) ~dA, \label{eq:2D_Model:BilinearChp5}\\
	c(\partial_t^2 u,\phi) & = & \int_\Omega~ (\partial^2_t u) \cdot \phi~dA, \label{eq:2D_Model:Bilinear_cChp5}\\
	(f,g) &=& \int_{\Omega} f\cdot g \ dA, \label{eq:2D_Model:Bilinear_intChp5}
\end{eqnarray}
with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1-\nu^2)}$.

Define the test function space 
\begin{eqnarray*}
	T(\Omega) & = & \left\{ \phi \in C^1(\bar{\Omega}) \ | \ \phi = 0 \ \textrm{ on } \ \Gamma \right\}.
\end{eqnarray*}

Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and 
\begin{align}
	c(u,\phi) = -b(u,\phi) + (Q,\phi) \label{eq:2D_Model:Problem2D1VEqChp5}
\end{align}
for all $\phi \in T(\Omega)$.

@@

\section{Cantilever three-dimensional model in variational form} \label{ssec:3D_Model:FEM}
	Let $\{e_1,e_2,e_3\}$ be a right-handed orthonormal basis for $R^3$. Denote the elastic body as $\Omega \in R^3$ with $(0,0,0)$ the point of reference. For a rectangular cross-section, the body $\Omega$ can be described as
	\begin{eqnarray*}
		\Omega = \left\{ x \in R^3 \ | \ 0 \leq x_1 \leq 1, \ -\frac{h}{2} \leq x_2 \leq \frac{h}{2} , \ -\frac{b}{2} \leq x_3 \leq \frac{b}{2}\right \}
	\end{eqnarray*}
	$\partial \Omega$ denotes the boundary of the body. $\Sigma$ is the part of the boundary where $x_1 = 0$ and here the boundary condition is $u = 0$. $\Gamma$ is the remainder of the boundary where $Tn = 0$.

@@

	\subsubsection*{Problem 3DV}
	The bilinear forms and integral function is defined by
	\begin{eqnarray}
		b(u,\phi) &=& \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
		\textrm{Tr}(\Phi) ~dV,\\
		c(u,\phi) &=& \int_\Omega~ (\partial^2_t u) \cdot \phi~dV, \\
		(f,g) &=& \int_{\Omega} f\cdot g \ dV, \\
	\end{eqnarray}
	with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1+\nu)(1-2\nu)}$.\\

	The test function space is defined as
	\begin{eqnarray*}
		T(\Omega) & = & \left\{ \phi \in C(\Omega) \ | \ \phi = 0 \ \textrm{ on } \ \Sigma \right\}.
	\end{eqnarray*}

	Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and 
	\begin{align}
		c(u,\phi) = -b(u,\phi) - (Q,\phi) \label{3DB_20}
	\end{align}
	for all $\phi \in T(\Omega)$.\\

@@

	\subsubsection*{Problem 3DG}
	Find a function $u^h$ such that for all $t>0$, $u^h \in S^h$ and
	\begin{eqnarray*}
		(u^h, \phi_i) & = & -b(u^h,\phi_i) + (Q^I \cdot \phi_i)
	\end{eqnarray*} for $i = 1,2,...,k$. $Q^I$ is scalar vector obtained after interpolating the function $Q$ over the brickable grid $\Omega$. i.e. $Q^I_{i,j, h} = Q(x_i,x_j, x_h)$ for $i = 1,2,...,n_1$, $j = 1,2,...,n_2$ and $h = 1,2,...,n_3$.

@@

\section{Validity of the model for a cantilever Timoshenko beam} \label{sec:validity-of-a-cantilever-timoshenko-beam}

	In this section, the validity of a cantilever Timoshenko beam model is investigated. This investigation is done in \cite{LVV09}, and is included here with more significant digits and some extended results.

	Recall that the cantilever Timoshenko beam model is referred to as Problem T-2 and the two-dimensional cantilever beam model is referred to as Problem 2D as defined in Section \ref{sec:LinearMathematicalModels}. Figure \ref{fig:compare:1D+2D} shows the two beams side-by-side.

@@

						%\node at (3.2,-0.2) {$\ell = 1$};
						\end{tikzpicture}
					\end{center}
					\subcaption{Timoshenko Cantilever Beam (Problem T-2)}
				\end{minipage}
				\begin{minipage}[b]{0.8\linewidth}
					\begin{center}
						\begin{tikzpicture}
							\draw[line width = 0.4mm] (0,0.2) -- (6,0.2);
							\draw[line width = 0.4mm] (0,-0.2) -- (6,-0.2);
							\draw[line width = 0.4mm] (6,-0.2) -- (6,0.2);
							%\draw[line width = 0.4mm] (0,-0.2) -- (0,0.2);

@@

						\end{tikzpicture}
					\end{center}
					\subcaption{Two-Dimensional Cantilever Beam (Problem 2D)}
				\end{minipage}
			}
		}
	\end{figure}\label{fig:compare:1D+2D}

	In introduction of the Timoshenko beam model, the parameter $\alpha$ is introduced. The paramater is given here again for convenience:
	\begin{eqnarray*}
		\alpha = \frac{A\ell^2}{I}.
	\end{eqnarray*}
